\chapter{\hlc[yellow]{Glossary}} \label{glossary}

Language in general and technical jargon (of which this glossary qualifies) in particular is intended to communicate. This requires that both the speaker and the listener have some common understanding of the terms used. For this reason, I rarely find it helpful to generate new definitions for commonly used words, except to clarify when there is some significant discrepancies in how the term is commonly used. It is generally preferable to coin a new term if a new meaning is required (see, for instance Myoa Bailey's coining of the term \textit{misogynoir} \cite{baileyMoreOriginMisogynoir} or the significantly less elegant \textit{socio-environmental-technical system} used in this document).  

\textbf{Accessibility:} The degree to which something (usually data in this thesis) is easy to access for a given person or group. Greater accessibility means that a larger set of people can access and use the data with minimal difficulty. To be compared with \textit{availability}.

\textbf{Availability:} The degree to which something (usually data in this thesis) is available for use by someone. Barriers to availability include a time delay and lack of sensors. Availability is primarily based on what data is ``out there" independent of how easy it is to gain access to it or to use it. To be compared with \textit{accessibility}. 

%\textbf{Collaborative Systems:} A system that is not under central control, either in its conception, development, or operation. They tend to be assembled and operated through the voluntary choices of the participants, not through the dictates of an individual client \cite{maierArtSystemsArchitecting2009}.

\textbf{Context:} See \textit{System Context}.

\textbf{Critical Remote Sensing:} A new field, most clearly laid out by Bennet et al. that reconsiders the rationales for the use of satellite data in a more critical light than has been common \cite{bennettPoliticsPixelsReview2022}. In particular, they advocate for a tripartite research agenda of \textit{exposing}, \textit{engaging}, and \textit{empowering}. By exposing, they mean using remote sensing to provide evidence of socioeconomic and environmental injustices, with a particular emphasis on clandestine activities. By engaging, they mean recognizing the very much non-objective perspective of remote sensing and seeking to integrate remote sensing with local knowledge rather than supplant it. By empowering, they mean partnering with groups that remote sensing is collecting data about, particularly marginalized groups, for capacity building and participating in the use of the data.

\textbf{\acf{dss}:} A technical system aimed at facilitating and improving decision-making. Functions can include visualization of data, analysis of past data, simulations of future outcomes, and comparisons of options.

\textbf{\acf{eo}:} As defined by \ac{geo}, the data and information collected about our planet, whether atmospheric, oceanic or terrestrial. This includes space-based or remotely-sensed data, as well as ground-based or in situ data \cite{grouponearthobservationsGEOGlance2019}. As defined by Mather and Koch, the interpretation and understanding of measurements of the Earth's land, ocean, or ice surfaces or within the atmosphere, together with the establishment of relationships between the measurements and the nature and distribution of phenomena on the Earth's surface or within the atmosphere \cite{matherComputerProcessingRemotelySensed2011}. Either definition is applicable, though typically \ac{eo} in this work refers to data collected via remote observation.

\textbf{Ecosystem Services:} The various benefits that humans are provided by the natural environment and healthy ecosystems in particular. Ecosystem services are often sorted into three categories: provisioning (providing some raw material), regulating (moderating the ambient environment in a helpful manner), and cultural (non-material benefits) \cite{haines-youngCommonInternationalClassification2018}. See Sections \ref{sec:sustain}, \ref{sec:sdg_critique}, and \ref{sec:rio-vulnerability} for more detailed discussion.

\textbf{\acf{evdt}:} A four-part modeling framework created by Space Enabled for use in \acp{sets} and sustainable development applications \cite{reidCombiningSocialEnvironmental2019}. For more detail, including diagrams, see Chapter \ref{ch:evdt}.

\textbf{Form:} See \textit{System Form}.

\textbf{Function:} See \textit{System Function}.

\textbf{\acf{gis}:} Any digital system for storing, visualizing, and analyzing geospatial data, that is data that has some geographic component. The term can also be used to discuss specific systems, a method that uses such systems, a field of studying focusing on or involving such systems, or even the set of institutions and social practices that make use of such a system \cite{sheppardGISSocietyResearch1995}. For more discussion of this definition, see Section \ref{sec:gis}.

\textbf{Multidisciplinary Optimization:} A methodology for the design of systems in which strong interaction between disciplines motivates designers to simultaneously manipulate variables in several disciplines \cite{sobieszczanski-sobieskiMultidisciplinaryAerospaceDesign1997}.

\textbf{Multi-Stakeholder Decision-Making:} Any decision-making process in which more than one stakeholder must collaborate to reach a decision \cite{fitzgeraldRecommendationsFramingMultistakeholder2016}. This can take a variety of forms, including cooperation, negotiation, voting, or consultation \cite{garberMultiStakeholderTradeSpace2015}.

\textbf{Objective:} See \textit{System Objective}.

\textbf{\acf{osse}:} A method of investigating the potential impacts of prospective observing systems through the generation of simulated observations that are then ingested into a data assimilation system and compared to other real-world data or other simulated data. Most commonly used for remote observation satellite design for purposes of meteorology \cite{masutaniObservingSystemSimulation2010} .

%\textbf{Organizational Policy:} Policy, decision-making, and politics within an organizational stakeholder. This includes decision-making policies, mechanisms of institutional learning and memory, capability development, etc. See \textcolor{black}{the Day 1 Response} for further discussion.

\textbf{\acf{pgis}:} A subset of \ac{gis} that seeks to directly involve the public and other stakeholders, including government officials, \acp{ngo}, private corporations, etc \cite{sieberPublicParticipationGeographic2006}. It should be noted that these means involvement in both the production of data and in its application, not merely one or the other \cite{weinerParticipatoryGeographicInformation2007, talenBottomUpGIS2000}. This is to be contrasted with the older term, \ac{ppgis}, which focuses specifically on the involvement of the public and not that of government agencies or other organizations \cite{sieberPublicParticipationGeographic2006}. For more discussion of this and related terms, see Section \ref{sec:collaborative}.

\textbf{Planning}: ``the premeditation of action, in contrast to management [which is] the direct control of action" \cite{harrisLocationalModelsGeographic1993}. In general, planning tends to concern itself with more long-term affairs that management does, during which it strives for the "avoidance of unintended consequences while pursuing intended goals." Models, and their specific implementations as decision/planning support tools, are one means of achieving this. The term is often prefaced with `urban' or `regional' to indicate the specific spatial scale under consideration.

\textbf{Primary Stakeholder:} Those who make direct decision on the design or operation of the system.

\textbf{\acf{pss}:} A type of \ac{dss} specifically designed to support urban or regional planning efforts. These often involve longer time scales and more general/strategic decisions than most \acp{dss}. In general, this work will use the more general term, \ac{dss}, and will only use \ac{pss} when referring to the literature.

\textbf{Remote Observation}: Any form of data collection that takes place at some remote distance from the subject matter \cite{jensenRemoteSensingEnvironment2006}. While there is no specific distance determining whether a collector is `remote,' in practice this tends to mean some distance of more than a quarter of a kilometer. Handheld infrared measurement devices are thus usually excluded (and thereby classified as \textit{in-situ} observations. Aerial and satellite imagery are definitively in the remote observation category. Low altitude drone imagery, particularly when the operator is standing in the field of view, is a gray area that is not well categorized at this time.

\textbf{Remote Sensing:} See \textit{remote observation}.

\textbf{Scenario Planning}: A particular form of planning that focuses on long-term strategic decisions through the representation of multiple, plausible futures of a system of interest \cite{goodspeedScenarioPlanningCities2020}. These futures are often generated by models such as \ac{evdt}.

\textbf{Secondary Stakeholder:} Those who have direct influence on the Primary Stakeholders, typically via authority or funding.

\textbf{Socio-environmental System:} The complex phenomena that occurs due to the interactions of human and natural systems \cite{elsawahEightGrandChallenges2020}.

\textbf{Sociotechnical System:} Technical works involving significant social participation, interests, and concerns \cite{maierArtSystemsArchitecting2009}.

\textbf{Socio-environmental-technical System:} A system in which social, environmental, and technical subsystems are linked together in such a way that none can be neglected without compromising the modeling, planning, or forecasting objectives at hand. This can be seen as the combination of the terms sociotechnical system and socio-environmental system. Note the particular emphasis on the needs of the observer, not the inherent system itself, as virtually all systems on Earth can be viewed as socio-environmental-technical Systems.

\textbf{Stakeholder:} The people, organizations, and communities that either influence the design and operation of the system or are impacted by the system. For more discussion, including alternative definitions, see Sections \ref{sec:saf} and \ref{sec:saf_stakeholders}.

\textbf{Stakeholder Analysis:} Identifying, mapping, and analyzing the stakeholders in a system and their connections to one another in order to inform the design of the system. This involves both qualitative and quantitative tools, such as the Stakeholder Requirements Definition Process \cite{incoseINCOSESystemsEngineering2015} and Stakeholder Value Network Analysis \cite{fengDependencyStructureMatrix2010a}. It should be noted that this term is commonly used by systems engineers but is not clearly defined as some specific list of methods. In a Space Enabled context, it commonly refers to the coding of qualitative interviews with stakeholders to elicit such items as needs, desired outcomes, and objectives. These are then often analyzed in some other method, such as Stakeholder Value Network Analysis. For more discussion, see Section \ref{sec:saf_stakeholders}.

\textbf{Sustainable Development}: The integration of three separate, previously separate fields: economic development, social development and environmental protection \cite{worldsummitonsustainabledevelopmentPlanImplementationWorld2002}.  For a more detailed discussion of the history of this term, see Section \ref{sec:sustain}.


\textbf{Systems Architecture/Architeting:} The mapping of function to form such that the essential features of the system are represented. The intent of architecture is to reduce ambiguity, employ creativity, and manage complexity \cite{crawleySystemArchitectureStrategy2015}. For more discussion including alternative definitions, see Section \ref{sec:saf_stakeholders}.

\textbf{System Boundary:} The limit demarcating the system of interest from the rest of the universe. The System Boundary determines what is considered part of the system (and thus subject to the decisions of the designer and stakeholders) and what is not (and is thus considered beyond their control, for the purposes of this project at least). See Section \ref{sec:saf_context} for more details.

\textbf{System Context:} The external factors that influence and constrain the system. For more discussion, see Section \ref{sec:saf_context}.

\textbf{Systems Engineering:} An interdisciplinary approach and means to enable the realization of successful systems. It focuses on holistically and concurrently understanding stakeholder needs; exploring opportunities; documenting requirements; and synthesizing, verifying, validating, and evolving solutions while considering the complete problem, from system concept exploration through system disposal \cite{systemsengineeringbodyofknowledgeSystemsEngineeringGlossary2021}. 
For a more detailed discussion of this definition, including its flaws, see Section \ref{sec:se}.

\textbf{System Form:} The approaches and structures used to enable the System Functions (i.e. the physical ``stuff" of the system).

\textbf{System Function:} The specific actions or processes that the system performs, with a particular emphasis on those in service of the System Objectives.

\textbf{System Objective:} The high level description of what the system will do.

\textbf{Tertiary Stakeholder:} Those that exert either little control or primarily indirect control on the system, but ae impacted by the system.

\textbf{Tradespace:}  The space spanned by the completely enumerated design variables, i.e. the set of possible design options \cite{rossTradespaceExplorationParadigm2005}.

\textbf{Tradespace Exploration:} A process by which various options with a tradespace may be examined and compared in the absence of a single utility function, such as when multiple stakeholders are involved or multiple contexts with no clear priority exist \cite{rossTradespaceExplorationParadigm2005}.

\clearpage
\newpage
