\chapter{\hlc[green]{Rio de Janeiro Stakeholder Interview Questions}} \label{interview-questions}

The following questions were used during the stakeholder interviews and meetings conducted during field visits to Rio de Janeiro in August of 2019 and March of 2020. This list does not include followup questions triggered by stakeholder responses.

\begin{enumerate}\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
    \item{To confirm, are you okay being recorded?}
    \item{What is your name?}
    \item{What organization are you associated/work with?}
    \item{What is your role there?}
    \item{What is your organization's primary goal/mission? How does it usually pursue that goal?}
    \item{What are some example projects/activities?}
    \item{How do individual projects in your organization originate?}
    \item{What stakeholders does your organization work with?}
    \item{What do you view as the primary pressures on mangroves in the Rio de Janeiro area?}
    \item{What do you view as the primary pressures on the people living near the mangroves?}
    \item{What, if any, interest does your organization have in the Rio de Janeiro mangroves?}
    \item{Does your organization use any remote sensing data?}
    \item{Has your organization ever participated in the design of an earth observation satellite? Or considered doing so?}
    \item{What other sources of data does your organization rely upon?}
    \item{What are some questions that you would like to be able to answer but can't? What are some challenges that your organization faces?}
    \item{How much are you or your organization able to explore new data sources or data analysis methods, as opposed to continuing to rely upon your current methods?}
    \item{Anything else you want to add?}
\end{enumerate}